---
title: "Exploratory time series analysis"
author: "Chiara Todaro"
date: "12 April 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation

Time series analysis is central tool in economics, physics, process and quality control, inventory studies, and many others fields. 

The starting point for a good analysis is to get a good understanding of the data by making some exploratory analyses.

This simple [app](https://chiara87todaro.shinyapps.io/EA_seatbelts/), built with the R 
package *shiny*, shows a few techniques that can be used to analyse data in R.

In the next slides a brief presentation of the data set and examples of the tecniques used is given.

Have fun!
 <!-- Exploratory data analysis is a key part of the data science process because it allows you to sharpen your question and refine your modeling strategies.
 R.Peng, Exploratory Data Analysis with R -->
 <!--  <http://rmarkdown.rstudio.com>. -->


## Data set
 
**Seatbelts** dataset is a multivariate time series collecting monthly information 
of serious car accidents happened in Great Britain between January 1969 and December 1984.

```{r data,collapse=TRUE,results='hold'}
Seatbelts[1:8,]
```


## Technique 1: moving average
<!-- **ma()** smooths data by averaging over sliding time-windows.  -->
<!-- It can be relaized with the function **ma()** in the R package *forecast*. -->
<!-- Let's see what happen to a gaussian random time series. -->
```{r ma, collapse = TRUE}
# smooths data by averaging over sliding time-windows
library(forecast)# x=ts object, order= length of time-window
mov_av<-ma(x=ts(rnorm(1000)),order=50);plot(mov_av)
```

## Technique 2: cross-correlation

```{r pressure}
# cross-correlation at different lags between two ts objects
ccf(x=ts(rnorm(1000)),y=ts(rnorm(1000)),lag.max=10,type="correlation")
```
